Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-08 article/article.Rmd
Commit 2024-06-08 09:12

<!-- Define CE and ECE -->
In the relative bibliography different definition have been used for these events,
some of the are summarized by @Gueymard2017.  Here, we defined as CE
events, the cases when the measured GHI at ground
level, exceeds the expected value under clear-sky conditions.  Similarly, we define as
extreme cloud enhancement events (ECE), the cases when GHI at ground level exceeds the Total Solar
Irradiance (TSI).  Although the duration of these
bursts can vary, from seconds to several minutes, here we are constrained by the
recorded data, to one-minute steps.

@@

The monitoring site is operating in the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
\(22^\circ\,57'\,\)E, \(80\,\)m\ a.s.l.).
In this study we present data from the period
`r strftime(min(ST_daily$Date), "%d\\ %B %Y")` to
`r strftime(max(ST_daily$Date), "%d\\ %B %Y")`.
<!-- Instrument -->
The GHI data were measured with a Kipp\ & Zonen CM-21 pyranometer.  During the study
period, the pyranometer was independently calibrated three times at the
Meteorologisches Observatorium Lindenberg, DWD, verifying that the stability of the
instrument’s sensitivity was better than $0.7\,\%$ relative to the initial
calibration by the manufacturer.
<!-- Measurements -->
For the acquisition of radiometric data, the signal of the pyranometer was sampled at
a rate of $1\,\text{Hz}$.  The mean and the standard deviation of these samples were
calculated and recorded for every minute.  The measurements were corrected for the
zero offset ("dark signal" in volts), which was calculated by averaging all
measurements recorded for a period of $3\,\text{h}$, before (morning) or after
(evening) the Sun reaches an elevation angle of $-10^\circ$.  The signal was
converted to irradiance using the ramped value of the instrument’s sensitivity
between subsequent calibrations.

<!-- Radiation data quality check -->
To further improve the quality of the irradiance data, a manual screening was
performed, in order to remove inconsistent and erroneous recordings that can occur
stochastically or systematically during the long operation of the instrument.
The manual screening was aided by a radiation data quality assurance procedure,
adjusted for the site, which was based on the methods of Long and Shi\ [@Long2006;
@Long2008a].  Thus, problematic recordings have been excluded from further
processing.  Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the quality of the
radiation measurements used in this study.
<!-- Data selection -->
To preserve an unbiased representation of the data we applied a constraint, similar
the one used by @CastillejoCuberos2020. Where, for each valid hour of day, there must
exist at least 45 minutes of valid measurements, including nighttime, when near sunrise and
sunset. Days with less than 5 valid hours were rejected completely.
<!--# CastillejoCuberos2020 There must be at least 5 h of valid data (Tests 1–7),
otherwise all data for this day is flagged as invalid. There must be at least This
work valid data 45 min of valid data (Tests 1–7) for an hour to be flagged as valid,
otherwise all minute data for this hour is flagged as invalid. -->
Furthermore, due to the significant measurement uncertainty, when the Sun is near the
horizon, and due to some systematic obstructions by nearby buildings, we have
excluded all measurements with solar zenith angle (SZA) greater than 
$`r 90 - BIO_ELEVA`^\circ$.

@@

To be able to detect the CE cases, we had to establish a baseline, above which we 
characterized each data point as an enhancement event, by estimating the occurring
over irradiance (OIR). The OIR, is defined as the irradiance difference of the
measured one-minute $\text{GHI}_i$ from the CE identification criterion
($\text{OIR}_i = \text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}.
To have an estimation, and a first insight on the phenomenon, we experimented with
two simple approaches for the reference.  The Haurwitz's model [@Haurwitz1945], which
is as simple clear sky model, and we had already adjusted and had good fit with our
data [@Natsis2023], and the TSI at the top of the atmosphere.  We have tested both
cases by using an appropriate relative threshold and/or an additional constant
offset.  The initial results, showed that we can detect a big portion of the CE
events. These results were helpful, as they are independent from unknown factors, and
helped us to establish some criteria to further improve the CE identification. It was
evident, by inspecting the daily plot of irradiance, that changes on the atmospheric
conditions introduced numerous false positive and false negative results.  To produce
a more accurate reference, we had to take into account more factors that effect the
clear sky radiation. So we used a radiation transfer
model in order to include the effects of ADO and the water vapors.

@@

<!-- Libradtran options -->
For completeness, we will describe here the main points of the radiation simulation
procedure. We used as input the spectrum of @Kurucz1994 in the range $280$ to
$2500\,\text{nm}$, with the Libradtran radiation transfer solver "disort" on a
"pseudospherical" geometry and the "LOWTRAN" gas parameterization. For each
combination of conditions we use a SZA step of $0.2^\circ$.
<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated on combinations of AOD at
$500\,\text{nm}$ ($\tau_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w$), also with offsets of $\pm1$ and $\pm2\sigma$.
We applied them on two atmospheric profiles, from the Air Force Geophysics
Laboratory (AFGL). The "AFGL atmospheric constituent profile, midlatitude summer"
(afglms) and the "AFGL atmospheric constituent profile, midlatitude winter" (afglmw)
[@Anderson1986].

<!-- Lookup table -->
To create a look-up table, that aligns with our dataset, we applied some adjustments.
To account for the Sun's variability, in our one-minute GHI measurements, we adapted
each modeled value by scaling the model's input spectrum integral, to the
corresponding TSI, provided by NOAA [@Coddington2005].  Also,
we account for the effect of the Earth -- Sun distance on the irradiance, by using
the distance calculated by the Astropy [@AstropyCollaboration2022] software library.
As needed, we interpolate the resulting irradiances to the SZA of our
measurements.  For each period of the year, we used the appropriate atmospheric
profile (afglms or afglmw).  Finally, we calculated the clear sky irradiance value at
the horizontal plane.  Thus, we were able to emulate different atmospheric condition
and levels of atmospheric clearness for the climatological conditions of the site.
With this method, the modeled clear sky irradiances can be directly compared to each
measured one-minute value of GHI, for different conditions of atmospheric clearness.

>>>>

<!-- Define CE and ECE tomove? -->
In this study, we define an event as CE when the measured GHI at ground level,
exceeds the expected value under clear-sky conditions. Similarly, we define as
extreme cloud enhancement events (ECE), the cases when GHI at ground level exceeds
the Total Solar Irradiance (TSI). Although the duration of these bursts can vary from
seconds to several minutes, here we are constrained by the temporal resolution of our
data to identify evens with duration of at minimum one-minute.

@@

The data used in this study were recorded at the monitoring site of the Laboratory of
Atmospheric Physics, Aristotle University of Thessaloniki, in Thessaloniki, Greece
(\(40^\circ\,38'\,\)N, \(22^\circ\,57'\,\)E, \(80\,\)m\ a.s.l.).  The GHI data were
measured with a Kipp\ & Zonen CM-21 pyranometer and cover the period
`r strftime(min(ST_daily$Date), "%d\\ %B %Y")` to
`r strftime(max(ST_daily$Date), "%d\\ %B %Y")`.
<!-- Instrument -->
  During the study
period, the pyranometer was independently calibrated three times at the
Meteorologisches Observatorium Lindenberg, DWD, verifying that the stability of the
instrument’s sensitivity was better than $0.7\,\%$ relative to the initial
calibration by the manufacturer.
<!-- Measurements -->
For the acquisition of radiometric data, the signal of the pyranometer was sampled at
a rate of $1\,\text{Hz}$ with the mean and standard deviation of these samples
calculated and recorded every minute.  The measurements were corrected for the zero
offset ("dark signal" in volts), which was calculated by averaging all measurements
recorded for a period of $3\,\text{h}$, before (morning) or after (evening) the Sun
reaches an elevation angle of $-10^\circ$.  The signal was converted to irradiance
using the ramped value of the instrument’s sensitivity between subsequent
calibrations.

<!-- Radiation data quality check -->
To further improve the quality of the irradiance data, a manual screening was
performed, in order to remove inconsistent and erroneous recordings that can occur
stochastically or systematically during the long operation of the instrument.  The
manual screening was aided by a radiation data quality assurance procedure, adjusted
for the site, which was based on the methods of Long and Shi\ [@Long2006;
@Long2008a].  Thus, problematic recordings have been excluded from further
processing.  Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the quality of the
radiation measurements used in this study.
<!-- Data selection -->
To preserve an unbiased representation of the data we applied a constraint, similar
the one used by @CastillejoCuberos2020. For each valid hour of day, there must exist
at least 45 minutes of valid measurements, including nighttime, when near sunrise and
sunset. Days with less than 5 valid hours were rejected completely.
<!--# CastillejoCuberos2020 There must be at least 5 h of valid data (Tests 1–7),
otherwise all data for this day is flagged as invalid. There must be at least This
work valid data 45 min of valid data (Tests 1–7) for an hour to be flagged as valid,
otherwise all minute data for this hour is flagged as invalid. -->
Furthermore, due to the significant measurement uncertainty in GHI when the Sun is
near the horizon, and due to some systematic obstructions by nearby buildings, we
have excluded all measurements with solar zenith angle (SZA) greater than $`r 90 -
BIO_ELEVA`^\circ$.

Finally, images from a sky camera have been used in the manual inspection of the CE
identification. The sky camera operates since 2012 and stores images in 5 min time
steps.

@@

For the detection of CE cases we established a baseline of irradiance above which we
characterized each data point as an enhancement event and calculated the over
irradiance (OIR). The OIR is defined as the irradiance difference of the measured
one-minute $\text{GHI}_i$ from the CE identification criterion ($\text{OIR}_i =
\text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}.
To have an estimation, and a first insight on the phenomenon, we experimented with
two simple approaches for the reference.  The Haurwitz's model [@Haurwitz1945], which
is as simple clear sky model, and we had already adjusted and had good fit with our
data [@Natsis2023], and the TSI at the top of the atmosphere.  We have tested both
cases by using an appropriate relative threshold and/or an additional constant
offset.  The initial results, showed that we can detect a big portion of the CE
events. These results were helpful, as they are independent from unknown factors, and
helped us to establish some criteria to further improve the CE identification. It was
evident, by inspecting the daily plot of irradiance, that changes on the atmospheric
conditions introduced numerous false positive and false negative results.  To produce
a more accurate reference, we had to take into account more factors that effect the
clear sky radiation. So we used a radiation transfer
model in order to include the effects of ADO and the water vapors.

@@

<!-- Libradtran options -->
<!-- 
For completeness, we will describe here the main points of the radiation simulation
procedure. We used as input the spectrum of @Kurucz1994 in the range $280$ to
$2500\,\text{nm}$, with the Libradtran radiation transfer solver "disort" on a
"pseudospherical" geometry and the "LOWTRAN" gas parameterization. For each
combination of conditions we use a SZA step of $0.2^\circ$.
<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated on combinations of AOD at
$500\,\text{nm}$ ($\tau_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w$), also with offsets of $\pm1$ and $\pm2\sigma$.
We applied them on two atmospheric profiles, from the Air Force Geophysics
Laboratory (AFGL). The "AFGL atmospheric constituent profile, midlatitude summer"
(afglms) and the "AFGL atmospheric constituent profile, midlatitude winter" (afglmw)
[@Anderson1986].
-->

<!-- Lookup table -->
<!-- 
To create a look-up table, that aligns with our dataset, we applied some adjustments.
To account for the Sun's variability, in our one-minute GHI measurements, we adapted
each modeled value by scaling the model's input spectrum integral, to the
corresponding TSI, provided by NOAA [@Coddington2005].  Also,
we account for the effect of the Earth -- Sun distance on the irradiance, by using
the distance calculated by the Astropy [@AstropyCollaboration2022] software library.
As needed, we interpolate the resulting irradiances to the SZA of our
measurements.  For each period of the year, we used the appropriate atmospheric
profile (afglms or afglmw).  Finally, we calculated the clear sky irradiance value at
the horizontal plane.  Thus, we were able to emulate different atmospheric condition
and levels of atmospheric clearness for the climatological conditions of the site.
With this method, the modeled clear sky irradiances can be directly compared to each
measured one-minute value of GHI, for different conditions of atmospheric clearness.
-->
